Run-Time Monitoring with Adjustable Overhead \\Using Dataflow-Guided Filtering

### 2015-01-06 evaluation.tex
Spacing around tables/figures.
Edits based on Ed's comments.

% Filter monitoring overheads
\begin{figure}
  \begin{center}
    %\includegraphics[width=\columnwidth, clip=true, trim=0 0.5in 0 0.1in]{figs/data_filtering.pdf}
    \includegraphics[width=\columnwidth]{figs/data_filtering.pdf}
    \vspace{-0.2in}
    \caption{Monitoring overhead with null metadata filtering.}
    \label{fig:evaluation.filtering}
    \vspace{-0.1in}
  \end{center}
\end{figure}

@@

% BC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_hb_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for BC.}
    \label{fig:evaluation.bc_sweep}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% UMC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for UMC.}
    \label{fig:evaluation.umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% LS sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_ls_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for LS.}
    \label{fig:evaluation.ls_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

% IMP sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true]{figs/data_imp_sweep.pdf}
    \vspace{-0.2in}
    \caption{Average error vs. varying overhead budget for instruction-mix profiling. Whisker bars show min and max errors.}
    \label{fig:evaluation.imp_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure*}

@@

We see that by varying the overhead budget, the coverage achieved also varies.
With only a 1.1x overhead target, array bounds check still
achieves over 80\% monitoring coverage on average. The high coverage
achieved with such low overhead is due to two main effects.  The first is that
monitoring can be done in parallel, providing a certain level of monitoring coverage without
introducing overhead. The second effect is that there may still exist a large number of
monitoring events that do not lead to a bounds checks. As a
result, dropping these events can reduce overhead without a large impact on
monitoring coverage.

@@

% Flex UMC sweep
\begin{figure}
  \begin{center}
    \vspace{-0.1in}
    \includegraphics[width=\linewidth]{figs/data_fpga_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for UMC running on an FPGA-based monitor.}
    \label{fig:evaluation.fpga_umc_sweep}
    \vspace{-0.1in}
  \end{center}
\end{figure}

@@

%In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
In this section, we evaluate the trade-offs between different dropping policies.
For sub-flow dropping, we identified sub-flow nodes through profiling. Because
of the need to profile the applications to determine sub-flow nodes, these
experiments were simulated for 1 billion instructions without fast-forwarding.
These results (not shown) indicated that sub-flow dropping produced similar
results in terms of coverage and overhead target matching to unrestricted
dropping for the benchmarks and monitors that we tested. Monitors and
benchmarks with different monitoring dependence graphs may see a more
pronounced trade-off between using unrestricted dropping and sub-flow dropping.

Since, sub-flow dropping showed little difference from unrestricted dropping,
we focus on comparing unrestricted dropping and source-only dropping. The
results shown in Figures~\ref{fig:evaluation.umc_policies} and
\ref{fig:evaluation.bc_policies} were run for 2 billion instructions after
fast-forwarding for 1 billion instructions.
The results are shown for an overhead target of 2.0x for UMC and 1.5x for BC.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overhead for UMC. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. For most benchmarks, we see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of short, independent monitoring dependence chains. Source-only dropping
causes overshoot of the overhead target for {\tt hmmer} and {\tt h264ref}.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that, in several cases, source-only dropping
achieves higher coverage than unrestricted dropping while still closely matching the overhead target. For example, {\tt perlbench} shows a 10\% increase in coverage and {\tt libquantum} shows an 18\% increase in coverage by using source-only dropping. This is due
to the fact that some of the overhead of monitoring for unrestricted dropping
is being spent on wasted work as discussed in Section~\ref{sec:policies.which}.

@@

Next, we evaluate these trade-offs between source-only dropping and unrestricted dropping for BC.
Figure~\ref{fig:evaluation.bc_exec_time} shows the overhead differences for BC
and Figure~\ref{fig:evaluation.bc_coverage} shows the coverage for BC.
From Figure~\ref{fig:evaluation.bc_exec_time}, we see that for several
benchmarks, source-only dropping fails to meet the specified overhead target.
The overshoot of the overhead target is quite high with overhead differences over 100\% for {\tt bzip2}, {\tt hmmer}, and {\tt h264ref}.
Since only infrequently occurring array allocations are considered as source events for BC, it can be 
difficult for source-only dropping to match overhead targets. Although
Figure~\ref{fig:evaluation.bc_coverage} shows higher coverage for
source-only dropping, this is largely due to the fact that it is running with
higher overhead.

@@

@@

% Multi-run coverage with unrestricted dropping
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_multirun_coverage.pdf}
    \vspace{-0.3in}
    \caption{Coverage over multiple runs for BC with a 10\% probability to not drop events.}
    \label{fig:evaluation.multirun}
    \vspace{-0.3in}
  \end{center}
\end{figure}

One application of partial monitoring with low overhead is to enable
cooperative debugging. 
The idea with cooperative debugging is to use partial
monitoring with very low overhead across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by using a probabilistic dropping policy.
Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check using
unrestricted, sub-flow, and source-only dropping policies with probabilistic dropping.
These numbers are averaged across all benchmarks. Here, we use a 10\%
probability
that events are not dropped.  Each run
was simulated for 500 million instructions. 
Since the goal is code coverage, coverage here is
measured as the percentage of static instructions which are monitored in at
least one of the runs.
We see that for the unrestricted dropping policy there is almost no increase in
coverage.  This is due to the fact that it is likely that at least one
monitoring event in a metadata dependence chain will be dropped with the unrestricted dropping
policy. 
Instead, sub-flow dropping and source-only
dropping are much better suited for achieving high coverage over multiple runs.
Source-only dropping shows a 6\% increase in coverage with only eight runs
while sub-flow dropping shows a 1\% increase in coverage over eight runs.
While sub-flow dropping shows a slower increase in coverage than source-only
dropping, sub-flow dropping is better able to meet overhead targets.
With enough runs, both sub-flow dropping and source-only dropping should be able to
reach 100\% code coverage.

@@

% Area and Power Overheads
\begin{table}[tb]
  \begin{center}
    \vspace{-0.0in}
    \begin{footnotesize}
    \input{tabs/area_power}
    \end{footnotesize}
    \caption{Average power overhead for dropping hardware. Percentages 
    are normalized to the main core power.}
    \vspace{-0.2in}
    \label{tab:evaluation.area_power}
  \end{center}
\end{table}

>>>>

% Filter monitoring overheads
\begin{figure}
  \begin{center}
    %\includegraphics[width=\columnwidth, clip=true, trim=0 0.5in 0 0.1in]{figs/data_filtering.pdf}
    \includegraphics[width=\columnwidth]{figs/data_filtering.pdf}
    \vspace{-0.2in}
    \caption{Monitoring overhead with null metadata filtering.}
    \label{fig:evaluation.filtering}
    \vspace{-0.3in}
  \end{center}
\end{figure}

@@

% BC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_hb_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for BC.}
    \label{fig:evaluation.bc_sweep}
    \vspace{-0.3in}
  \end{center}
\end{figure}

% UMC sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for UMC.}
    \label{fig:evaluation.umc_sweep}
    \vspace{-0.2in}
  \end{center}
\end{figure}

% LS sweep
\begin{figure}
  \begin{center}
    \includegraphics[width=\linewidth]{figs/data_ls_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage vs. varying overhead budget for LS.}
    \label{fig:evaluation.ls_sweep}
    \vspace{-0.3in}
  \end{center}
\end{figure}

% IMP sweep
\begin{figure*}
  \begin{center}
    \includegraphics[width=\linewidth, clip=true]{figs/data_imp_sweep.pdf}
    \vspace{-0.2in}
    \caption{Average error vs. varying overhead budget for instruction-mix profiling. Whisker bars show min and max errors.}
    \label{fig:evaluation.imp_sweep}
    \vspace{-0.3in}
  \end{center}
\end{figure*}

@@

We see that by varying the overhead budget, the coverage achieved also varies.
With only a 1.1x overhead target, array bounds check still
achieves over 80\% monitoring coverage on average. The high coverage
achieved with such low overhead is due to two main effects.  The first is that
monitoring can be done in parallel, providing a certain level of monitoring coverage without
introducing overhead. The second effect is that there may still exist a large number of
monitoring events that do not lead to bounds checks. As a
result, dropping these events can reduce overhead without a large impact on
monitoring coverage.

@@

% Flex UMC sweep
\begin{figure}
  \begin{center}
    \vspace{-0.1in}
    \includegraphics[width=\linewidth]{figs/data_fpga_umc_sweep.pdf}
    \vspace{-0.2in}
    \caption{Coverage versus varying overhead budget for UMC running on an FPGA-based monitor.}
    \label{fig:evaluation.fpga_umc_sweep}
    \vspace{-0.3in}
  \end{center}
\end{figure}

@@

%In this section we evaluate the trade-offs between unrestricted dropping and source-only dropping.
In this section, we evaluate the trade-offs between different dropping policies.
Figure~\ref{fig:evaluation.umc_exec_time} shows the difference between the
overhead budget and the run-time monitoring overhead for UMC when the overhead target is set to 2.0x. A positive value
means that the overhead target was overshot while a negative value indicates
that the overhead budget was met. For most benchmarks, we see similar results for unrestricted
dropping and source-only dropping due to the fact that UMC consists of a large
number of short, independent monitoring dependence chains. Source-only dropping
causes overshoot of the overhead target for {\tt hmmer} and {\tt h264ref}.
Figure~\ref{fig:evaluation.umc_coverage} shows the coverage of UMC for
unrestricted dropping and source-only dropping. We see that, in several cases, source-only dropping
achieves higher coverage than unrestricted dropping while still closely matching the overhead target. For example, {\tt perlbench} shows a 10\% increase in coverage and {\tt libquantum} shows an 18\% increase in coverage by using source-only dropping. This is due
to the fact that some of the overhead of monitoring for unrestricted dropping
is being spent on wasted work as discussed in Section~\ref{sec:policies.which}.

@@

Next, we evaluate these trade-offs between source-only dropping and unrestricted dropping for BC.
Figure~\ref{fig:evaluation.bc_exec_time} shows the overhead differences for BC
and Figure~\ref{fig:evaluation.bc_coverage} shows the coverage for BC.
Here, the overhead target is 1.5x.
From Figure~\ref{fig:evaluation.bc_exec_time}, we see that for several
benchmarks, source-only dropping fails to meet the specified overhead target.
The overshoot of the overhead target is quite high with overhead differences over 100\% for {\tt bzip2}, {\tt hmmer}, and {\tt h264ref}.
Since only infrequently occurring array allocations are considered as source events for BC, it can be 
difficult for source-only dropping to match overhead targets. Although
Figure~\ref{fig:evaluation.bc_coverage} shows higher coverage for
source-only dropping, this is largely due to the fact that it is running with
higher overhead.

@@

The results for the sub-flow dropping policy are not included in these graphs
because our profiling infrastructure to identify sub-flow nodes currently does
not support fast-forwarding. Instead, we compared the sub-flow dropping with
other policies by running simulations for 1 billion instructions without
fast-forwarding. The results (not shown) showed that sub-flow dropping produced
similar coverage and overhead target matching to unrestricted dropping for the
benchmarks and monitors that we tested.

@@

% Multi-run coverage with unrestricted dropping
\begin{figure}
  \begin{center}
    \includegraphics[width=\columnwidth]{figs/data_multirun_coverage.pdf}
    \vspace{-0.2in}
    \caption{Coverage over multiple runs for BC with a 10\% probability to not drop events.}
    \label{fig:evaluation.multirun}
    \vspace{-0.2in}
  \end{center}
\end{figure}

One application of partial monitoring with low overhead is to enable
cooperative debugging. 
The idea with cooperative debugging is to use partial
monitoring with very low overhead across a large number of users or runs. By
varying the pattern of partial monitoring done on each run, the goal is to
achieve high coverage across multiple runs. Varying the monitoring that is
done for different runs can be achieved by using a probabilistic dropping policy.
Figure~\ref{fig:evaluation.multirun} shows the total
coverage achieved over multiple runs for array bounds check using
unrestricted, sub-flow, and source-only dropping policies with probabilistic dropping.
These numbers are averaged across all benchmarks. Here, we use a 10\%
probability
that events are not dropped.  Each run
was simulated for 500 million instructions. 
Since the effectiveness of cooperative debugging is often measured by code coverage, coverage here is
measured as the percentage of static instructions which are monitored in at
least one of the runs.
We see that for the unrestricted dropping policy there is almost no increase in
coverage.  This is due to the fact that it is likely that at least one
monitoring event in a metadata dependence chain will be dropped with the unrestricted dropping
policy. 
Instead, sub-flow dropping and source-only
dropping are much better suited for achieving high coverage over multiple runs.
Source-only dropping shows a 6\% increase in coverage with only eight runs
while sub-flow dropping shows a 1\% increase in coverage over eight runs.
While sub-flow dropping shows a slower increase in coverage than source-only
dropping, sub-flow dropping is better able to meet overhead targets.
With enough runs, both sub-flow dropping and source-only dropping should be able to
reach 100\% code coverage.

@@

% Area and Power Overheads
\begin{table}[tb]
  \begin{center}
    \vspace{-0.0in}
    \begin{footnotesize}
    \input{tabs/area_power}
    \end{footnotesize}
    \caption{Average power overhead for dropping hardware. Percentages 
    are normalized to the main core power.}
    \vspace{-0.1in}
    \label{tab:evaluation.area_power}
  \end{center}
\end{table}
